---
title: "Simulations"
author: "Hadrien Lorenzo"
date: "2023-10-06"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load package **SEPaLS**, simulation functions and results

```{r load}
library(SEPaLS)
source("R/functions.R")
load("data/data_statists.RData") ## Load saved results
```

*Remark*: All the results can be retrieved using the corresponding chuncks.

# Set workspace

```{r set}
## Dimension & sample size
# Two dimension
p1 <- 3
p2 <- 30
# Selected dimension : p1 or p2
p <- p2
# Sample size
n <- 500

## Beta
beta <- c(rep(1,2),rep(0,p-2))/sqrt(2)

## Link function power
c_s<-c(1,0.5,0.25)

## Noise
# Standard deviation of epsilon
sigma <- 0.9*diag(p)
# Noise signal ratio
r <- 5

## Pareto parameters for Y distribution
# pareto.params[1] : Pareto index and also equal to 1/gamma where gamma is the tail index
# pareto.params[2] :  minimum possible value of Y
pareto.params <- c(5,2)

## Selected distribution : "pareto" or "student"
dist <- "pareto"

## copula. fam   : 
#    0= Independent copula
#    1 = Gaussian copula
#    5 = Frank copula
#    3 = Clayton copula
copula.fam <- 3

## Number of replications
N<-1000

## Kendalls tau parameters parameter 
thetas <- c(8,1/2,1/2,8)
signe_taus <- c(-1,-1,1,1)
taus <- c(-0.8,-0.2,0.2,0.8)

## Sparse regularization parameters
lambda <- c(0,1e-4,5e-4,1e-3)
n_lambdas <- length(lambda)

## vMF regularization parameters
kappa0 <- c(0,1e-4,3e-3,1e-2)
n_kappas <- length(kappa0)
mu0_1 <- beta
N_prior_NON_Null <- 15
mu0_2 <- c(rep(1,N_prior_NON_Null),rep(0,p-N_prior_NON_Null))/sqrt(N_prior_NON_Null)


# The 200 largest observations
k.threshold<- n - 100

# Visualization parameters
colors <- viridis::viridis(4)
alpha_transpa <- 0.15
```

# Simulations for von Mises-Fisher prior

## Start simulations

```{r vMF,eval=FALSE,include=TRUE}
results_1 = results_2 <- list()
for(i_t in 1:length(thetas)){
  set.seed(i_t)
  results_1[[i_t]] <- simu_process(mu0 = mu0_1,kappa0 = kappa0,n,p,beta,
                                   c_s,sigma,r,dist,
                                   pareto.params,
                                   copula.fam = copula.fam,
                                   copula.param = thetas[i_t],
                                   N,
                                   k.threshold,type="vMF",
                                   signe_tau = signe_taus[i_t])
  results_2[[i_t]] <- simu_process(mu0 = mu0_2,kappa0 = kappa0,n,p,beta,
                                   c_s,sigma,r,dist,
                                   pareto.params,
                                   copula.fam = copula.fam,
                                   copula.param = thetas[i_t],
                                   N,
                                   k.threshold,type="vMF",
                                   signe_tau = signe_taus[i_t])
}
```


## Observe results

Use the following function

```{r functvMF}
observe_vMF <- function(j){
  for(i_t in 1:length(thetas)){
    cos.rep_1 <- data.frame(results_1[[i_t]])
    cos.rep_2 <- data.frame(results_2[[i_t]])
    oo_beta_1=oo_beta_2 <- NULL
    file_1 <- paste("../SEPaLS_simulations/results/vMF/simu_vMF_c",j,
                    "_Theta",i_t,
                    "_mu1.pdf",sep="")
    file_2 <- paste("../SEPaLS_simulations/results/vMF/simu_vMF_c",j,
                    "_Theta",i_t,
                    "_mu2.pdf",sep="")
    for(i in 1:n_lambdas){
      id <- which(cos.rep_1$kappa0==i &
                    cos.rep_1$id_power==j)
      if(i==1){
        oo_beta_1_median <- cos.rep_1$median[id]
        oo_beta_2_median <- cos.rep_2$median[id]
        oo_beta_1_quant5 <- cos.rep_1$quant5[id]
        oo_beta_2_quant5 <- cos.rep_2$quant5[id]
        oo_beta_1_quant95 <- cos.rep_1$quant95[id]
        oo_beta_2_quant95 <- cos.rep_2$quant95[id]
      }
      else
      {
        oo_beta_1_median <- cbind(oo_beta_1_median,
                                  cos.rep_1$median[id])
        oo_beta_2_median <- cbind(oo_beta_2_median,
                                  cos.rep_2$median[id])
        oo_beta_1_quant5 <- cbind(oo_beta_1_quant5,
                                  cos.rep_1$quant5[id])
        oo_beta_2_quant5 <- cbind(oo_beta_2_quant5,
                                  cos.rep_2$quant5[id])
        oo_beta_1_quant95 <- cbind(oo_beta_1_quant95,
                                  cos.rep_1$quant95[id])
        oo_beta_2_quant95 <- cbind(oo_beta_2_quant95,
                                  cos.rep_2$quant95[id])
      }
    }
    # pdf(file_1,width = 7,height = 5,onefile = T)
    main <- bquote(mu[0]==beta~", "~tau==.(taus[i_t])~" and "~c==.(c_s[j]))
    XX <- n-unique(cos.rep_1$nb_exceed)
    matplot(XX,oo_beta_1_median,
            type="l",lty=1,lwd=3,main=main,
            ylab=bquote("PC(Y"["n-k+1,n"]~")"),
            xlab=expression(k),
            ylim=c(0,1),
            col=colors)
    for(jj in 1:ncol(oo_beta_1_quant5)){
      polygon(x=c(XX,rev(XX)),
              y=c(oo_beta_1_quant5[,jj],rev(oo_beta_1_quant95[,jj])),
              col=scales::alpha(colors[jj],alpha_transpa),border=NA)
    }
    abline(h=(0:5)/5,col="gray",lty=3)
    abline(v=(0:10)*20,col="gray",lty=3)
    # dev.off()
    # pdf(file_2,width = 7,height = 5,onefile = T)
    main <- bquote(mu[0]==tilde(beta)~", "~tau==.(taus[i_t])~" and "~c==.(c_s[j]))
    matplot(n-unique(cos.rep_2$nb_exceed),oo_beta_2_median,
            type="l",lty=1,lwd=3,main=main,
            ylab=bquote("PC(Y"["n-k+1,n"]~")"),
            xlab=expression(k),
            ylim=c(0,1),
            col=colors)
    for(jj in 1:ncol(oo_beta_2_quant5)){
      polygon(x=c(XX,rev(XX)),
              y=c(oo_beta_2_quant5[,jj],rev(oo_beta_2_quant95[,jj])),
              col=scales::alpha(colors[jj],alpha_transpa),border=NA)
    }
    abline(h=(0:5)/5,col="gray",lty=3)
    abline(v=(0:10)*20,col="gray",lty=3)
    # dev.off()
  }
}
```

### For $c=1$
```{r obs_vMF1}
observe_vMF(which(c_s==1))
```

### For $c=1/2$
```{r obs_vMF2}
observe_vMF(which(c_s==1/2))
```

### For $c=1/4$
```{r obs_vMF3}
observe_vMF(which(c_s==1/4))
```


# Simulations for Sparse prior

```{r paraSimu}
p_1 <- 30
p_2 <- 300
beta_30 <- c(rep(1,2),rep(0,p_1-2))/sqrt(2)
beta_300 <- c(rep(1,2),rep(0,p_2-2))/sqrt(2)
```

## Start simulations

```{r sparse,eval=FALSE,include=TRUE}
results_sparse_p_1 = results_sparse_p_2 <- list()
for(i_t in 1:length(thetas)){
  set.seed(i_t)
  results_sparse_p_1[[i_t]] <- simu_process(mu0 = mu0_1,kappa0 = lambda,n,p_1,beta_30,
                                     c_s,sigma,r,dist,
                                     pareto.params,
                                     copula.fam = copula.fam,
                                     copula.param = thetas[i_t],
                                     N,
                                     k.threshold,type="Laplace",
                                     signe_tau = signe_taus[i_t])
  results_sparse_p_2[[i_t]] <- simu_process(mu0 = mu0_2,kappa0 = lambda,n,p_2,beta_300,
                                     c_s,sigma,r,dist,
                                     pareto.params,
                                     copula.fam = copula.fam,
                                     copula.param = thetas[i_t],
                                     N,
                                     k.threshold,type="Laplace",
                                     signe_tau = signe_taus[i_t])
}
```

## Observe results

Use the following function

```{r functSparse}
observe_sparse <- function(j){
  for(i_t in 1:length(thetas)){
    cos.rep_1 <- data.frame(results_sparse_p_1[[i_t]])
    cos.rep_2 <- data.frame(results_sparse_p_2[[i_t]])
    oo_beta_1=oo_beta_2 <- NULL
    file_1 <- paste("../SEPaLS_simulations/results/Laplace/simu_Laplace_c",j,
                    "_Theta",i_t,
                    "_mu1.pdf",sep="")
    file_2 <- paste("../SEPaLS_simulations/results/Laplace/simu_Laplace_c",j,
                    "_Theta",i_t,
                    "_mu2.pdf",sep="")
    for(i in 1:n_lambdas){
      id <- which(cos.rep_1$lambda==i &
                    cos.rep_1$id_power==j)
      if(i==1){
        oo_beta_1_median <- cos.rep_1$median[id]
        oo_beta_2_median <- cos.rep_2$median[id]
        oo_beta_1_quant5 <- cos.rep_1$quant5[id]
        oo_beta_2_quant5 <- cos.rep_2$quant5[id]
        oo_beta_1_quant95 <- cos.rep_1$quant95[id]
        oo_beta_2_quant95 <- cos.rep_2$quant95[id]
      }
      else
      {
        oo_beta_1_median <- cbind(oo_beta_1_median,
                                  cos.rep_1$median[id])
        oo_beta_2_median <- cbind(oo_beta_2_median,
                                  cos.rep_2$median[id])
        oo_beta_1_quant5 <- cbind(oo_beta_1_quant5,
                                  cos.rep_1$quant5[id])
        oo_beta_2_quant5 <- cbind(oo_beta_2_quant5,
                                  cos.rep_2$quant5[id])
        oo_beta_1_quant95 <- cbind(oo_beta_1_quant95,
                                   cos.rep_1$quant95[id])
        oo_beta_2_quant95 <- cbind(oo_beta_2_quant95,
                                   cos.rep_2$quant95[id])
      }
    }
    # pdf(file_1,width = 7,height = 5,onefile = T)
    XX <- n-unique(cos.rep_1$nb_exceed)
    main <- bquote(d==30~", "~tau==.(taus[i_t])~" and "~c==.(c_s[j]))
    matplot(XX,oo_beta_1_median,
            type="l",lty=1,lwd=3,main=main,
            ylab=bquote("PC(Y"["n-k+1,n"]~")"),
            xlab=expression(k),
            ylim=c(0,1),
            col=colors)
    for(jj in 1:ncol(oo_beta_1_quant5)){
      polygon(x=c(XX,rev(XX)),
              y=c(oo_beta_1_quant5[,jj],rev(oo_beta_1_quant95[,jj])),
              col=scales::alpha(colors[jj],alpha_transpa),border=NA)
    }
    abline(h=(0:5)/5,col="gray",lty=3)
    abline(v=(0:10)*20,col="gray",lty=3)
    # dev.off()
    # pdf(file_2,width = 7,height = 5,onefile = T)
    main <- bquote(d==300~", "~tau==.(taus[i_t])~" and "~c==.(c_s[j]))
    matplot(XX,oo_beta_2_median,
            type="l",lty=1,lwd=3,main=main,
            ylab=bquote("PC(Y"["n-k+1,n"]~")"),
            xlab=expression(k),
            ylim=c(0,1),
            col=colors)
    for(jj in 1:ncol(oo_beta_1_quant5)){
      polygon(x=c(XX,rev(XX)),
              y=c(oo_beta_2_quant5[,jj],rev(oo_beta_2_quant95[,jj])),
              col=scales::alpha(colors[jj],alpha_transpa),border=NA)
    }
    abline(h=(0:5)/5,col="gray",lty=3)
    abline(v=(0:10)*20,col="gray",lty=3)
    # dev.off()
  }
}
```

### For $c=1$
```{r obs_sparse1}
observe_sparse(which(c_s==1))
```

### For $c=1/2$
```{r obs_sparse2}
observe_sparse(which(c_s==1/2))
```

### For $c=1/4$
```{r obs_sparse3}
observe_sparse(which(c_s==1/4))
```

